Title goes here

### 2016-10-23 paper/paper.tex
PAPER: Add content

\subsection{OpenSky Network}
Most of aircraft are provided with ADB-S transmitter, which send continuosly information about the status and the position of the aircraft to guarantee better air traffic control. The messages are not encripted end, in most region, this technology will become mandatory in 2020. For these reasons OpenSky network was recently born. \\
OpenSky is an open source network that aggregates ADS-B messages collected by volunteers around the globe. The dataset is accessible to everyone for non-profit purposes and the data is provided in the raw format, in order to perform research and statistics about air traffic. \\
However there are some limitation to the data. Only part of Europe has a good coverage, which means there are many areas where is not yet possible to detect airplanes informations. Furthermore messages are collected by volunteers and there is not guarantee that antennas are working properly 24/7.

@@

    The ADS-B standard defines different kinds of messages which are sent at different frequencies:
    \begin{itemize}
        \item \textit{Aircraft identification} messages contain the callsign of the airplane;
        \item \textit{Airborne position} messages contain altitude, latititude and longitude of the airplane and are sent twice per second;
        \item \textit{Airborne velocity} messages contain climb rate and velocity of the airplane.
    \end{itemize}
    In addition to these messages, other emergencies, priority, capability, navigation accuracy category, and operational modes messages could be broadcasted.

\subsection{Route detection}
A way to reduce the number of air routes is illustrated in \textit{J. DeArmon et al, Air Route Clustering for a Queuing Network Model of the National Airspace System}. The research showed that similar routes can be grouped to reduce the number of path between a pair of airports thus decrease air traffic congestion. The same approach could be used to detect standard routes where aircraft fly, given the departure and arrival airports. The first prototype of this model has been developed to improve the air traffic flow in the National Airspace System of the USA.

@@

\subsection{Algorithms}

\subsubsection{Positions}

@@

\subsubsection{Flights}

\paragraph{Identification of discrete flights}

@@

\paragraph{Departure and arrival airports}

@@

To detect the airports, we used the help of an external dataset containing a
list of airports together with their locations. This dataset originally
contained much more data than we needed, as it also listed small non-commercial
airports. In order for our algorithm to be reasonably correct, we had to trim
the dataset first removing the airports without a IATA code, and then keeping
only commercial airports with the help of an external list we found on the
Internet. This filtering reduced the list from XXX to XXX entries.
The resulting number of entries is still a little bit high for our purposes, but
it was nevertheless the best dataset we could get. The alternative would have
been to build a dataset by hand, which seemed extremely time consuming.
% TODO ^^^^^^^^^^

@@

\paragraph{Sectors}

@@

\subsubsection{Routes}

@@

\paragraph{Clustering algorithm}

@@

@@

  * What are the standard routes?

The results from the complete execution of the algorithm are quite good.
It should be considered that there are a lot of minor airports, and flight
information for those airports is, as one could expect, incomplete.
Of 1165 of the total combinations of airports, about 85 \% of them has their
standard route determined by under 10 flights. These pairs usually correspond
to minor airports, and the accuracy of the computed standard route has
necessarily to be considered pretty low. For major airports, however, the
available data rises significantly. As an example, the standard route between
Schiphol and London Gatwick and converse has been determined by 39,
respectively 42 routes.

% TODO say something about the distance function

  * Which flights and airlines diverge from them? In which airports?
  * How much CO2 would be saved if aircrafts followed a straight-line route
    instead of the standard one?

* Is it possible to determine flights from raw adsb data?

@@

The next step would be to improve the algorithms, by making them more clever and
use more data from the dataset in a better way. The flight detection algorithm
could be improved by using an external dataset relating callsigns, which are
amost fully available from the data, to flights. At the time of writing we are
not aware of such dataset being publicly available, but there could be in the
future. Also, our flight detection algorithm does not take velocity messages
into account. Velocity information is not vital in flights detection, but we do
not exclude that the heading information of an aircraft could be combined with
its altitude information to better recognize ascents and descents.
The grouping criteria in our flight detection are quite restrictive,
to be able to get sufficiently reliable outputs with a relatively simple
algorithm. These criteria can surely be made less restrictive, using some more
advanced techniques to make up for missing points.

There is room for improvement in the route detection algorithm, for example
by tuning the way routes are clustered together, and in particular by selecting
a better distance function. Another improvement would also come naturally by
increasing the number of flights available to the analysis.
A different approach could be to discard the current algorithm and instead use
an external dataset of standard routes. In this setting, standard routes need
not be computed. Flights that diverge from them can simply be discovered by
computing the distance from the standard route for the particular pair of
airports, and checking the result against a threshold.

>>>>

\subsection{OpenSky Network}

Most of aircraft are provided with ADB-S transmitter, which send continuosly information about the status and the position of the aircraft to guarantee better air traffic control. The messages are not encripted end, in most region, this technology will become mandatory in 2020. For these reasons OpenSky network was recently born.

OpenSky is an open source network that aggregates ADS-B messages collected by volunteers around the globe. The dataset is accessible to everyone for non-profit purposes and the data is provided in the raw format, in order to perform research and statistics about air traffic. \\
However there are some limitation to the data. Only part of Europe has a good coverage, which means there are many areas where is not yet possible to detect airplanes informations. Furthermore messages are collected by volunteers and there is not guarantee that antennas are working properly 24/7.

@@

    The ADS-B standard defines different kinds of messages which are sent at different frequencies:

    \begin{itemize}
        \item \textit{Aircraft identification} messages contain the callsign of the airplane;
        \item \textit{Airborne position} messages contain altitude, latititude and longitude of the airplane and are sent twice per second;
        \item \textit{Airborne velocity} messages contain climb rate and velocity of the airplane.
    \end{itemize}

    In addition to these messages, other emergencies, priority, capability, navigation accuracy category, and operational modes messages could be broadcasted.

\subsection{Routes detection}

A way to reduce the number of air routes is illustrated in \textit{J. DeArmon et al, Air Route Clustering for a Queuing Network Model of the National Airspace System}. The research showed that similar routes can be grouped to reduce the number of path between a pair of airports thus decrease air traffic congestion. The same approach could be used to detect standard routes where aircraft fly, given the departure and arrival airports. The first prototype of this model has been developed to improve the air traffic flow in the National Airspace System of the USA.

    \subsubsection{Routes distance}

    Routes are represented by an ordered list of the geographical sectors where the aircrafts transit. Prior to define the clustering algorithm, it is important to define how to measure the distance between two different routes. Since each sector is identified by a code, the string editing distance can be used to measure the distance between two different routes. The distance is therefore represented by the minimum number of operation -- insertion, deletion and replacement -- to transform Route 1 into Route 2 or vice-versa. Different weight can also be applied to different operations. The distance should not depend by the length of the routes and for this reason it is then divided by the sum of the lengths of the routes. The final result represent the percentage of difference between the routes.

    \subsubsection{Routes clustering}

    Given a distance measure for each couple of routes, it is necessary to define a clustering algorithm to group them. Many methods has been tried during the research and the \textit{Leader algorithm} showed the best results. It proceed as follows:
    \begin{enumerate}
      \item A random route is assigned as leader of a new cluster;
      \item All the other routes which distance from the leader is less than a threshold are assigned to the cluster;
      \item Repeat the steps with the remaining routes until all routes are assigned to a cluster.
    \end{enumerate}

@@

\subsection{Algorithm overview}

Our processing algorithm is composed of three phases that are executed in
sequence. The output of a phase is the input of the next. The process starts
from a dataset of raw ADS-B messages encoded in AVRO format, and ends producing
the data necessary to answer our research questions. The phases are the
following:

\begin{enumerate}
  \item \textbf{Position extraction}: the dataset of AVRO-encoded ADS-B messages
  is analyzed to extract only the position information abount the aircrafts. In
  particular, this phase produces a table of timestamped position points with
  latitude, longitue and altitude. Every position point is associated to a
  particular aircraft, identified by a unique ICAO code.
  \item \textbf{Flight detection}: the unstructured position points are analyzed
  per aircraft (per ICAO), and grouped together to extract single discrete
  flights. For each detected flight, departure and arrival airports are
  computed.
  \item \textbf{Route detection}: flights are grouped by pair of departure and
  arrival airport, and sent to an aggregative clustering algorithm that is used
  to identify standard routes and routes that diverge from them.
\end{enumerate}

These three phases are explained in more detail in the sections that follow.

\subsection{Positions}

@@

\subsection{Flights}

@@

\subsubsection{Departure and arrival airports}

@@

To detect the airports, we used the help of an external dataset containing a
list of airports together with their locations. This dataset originally
contained much more data than we needed, as it also listed small non-commercial
airports. In order for our algorithm to be reasonably correct, we had to trim
the dataset first removing the airports without a IATA code, and then keeping
only commercial airports with the help of an external list we found on the
Internet. This filtering reduced the list from 6977 to 639 entries.
The resulting number of entries is still a little bit high for our purposes, but
it was nevertheless the best dataset we could get. The alternative would have
been to build a dataset by hand, which seemed extremely time consuming.

@@

\subsubsection{Sectors}

@@

\subsection{Routes}

@@

\subsubsection{Clustering algorithm}

@@

To aid the visualization, we consider ``diverging clusters'' instead of
diverging flights. After the clustering process, all clusters different from the
one selected as standard are diverging clusters. We then display the
representatives of the diverging clusters as the diverging routes for each
airport. The rationale is simply that routes in the same cluster are close
enough to be abstracted by a single representative of the entire group.

@@

The results from the complete execution of the algorithm are mixed.
It should be considered that there are a lot of minor airports, and flight
information for those airports is, as one could expect, incomplete.
Of 1165 of the total combinations of airports, about 85 \% of them has their
standard route determined by under 10 flights. These pairs usually correspond
to minor airports, and the accuracy of the computed standard route has
necessarily to be considered pretty low. For major airports, however, the
available data rises significantly. As an example, the standard route between
Schiphol and London Gatwick and converse has been determined by 39,
respectively 42 routes.

Tuning the distance function for the route clustering phase was more challanging
than we expected. In many cases, the algorithm ended up considering as diverging
routes that were actually very close to the standard one. We can conclude,
therefore, that even though we think we were able to determine standard routes
with acceptable accuracy and confidence for medium to big airports, diverging
routes detection still has some issues. We do think that the approach of using
sectors and editing distance can give good results in this context, but the
distance function obviously needs much more experiments are fine-tuning to be
able to give good results in all situations.

@@

The next step would be to improve the algorithms, by making them more clever and
use more data from the dataset in a better way. A trivial but effective
improvement to all phases of the algorithm follows, clearly, from increasing
the number of flights available to them.

The flight detection algorithm could be improved by using an external dataset
relating callsigns, which are amost fully available from the data, to flights.
At the time of writing we are not aware of such dataset being publicly
available, but there could be in the future. Also, our flight detection
algorithm does not take velocity messages into account. Velocity information is
not vital in flights detection, but we do not exclude that the heading
information of an aircraft could be combined with its altitude information to
better recognize ascents and descents. The grouping criteria in our flight
detection are quite restrictive, to be able to get sufficiently reliable outputs
with a relatively simple algorithm. These criteria can surely be made less
restrictive, using some more advanced techniques to make up for missing points.

There is room for improvement in the route detection algorithm, for example
by tuning the way routes are clustered together, and in particular by selecting
a better distance function.
A possible alternative way to compute the distance could be to compute the area
delimited by two routes. This area, normalized according to the length of the
routes, could be a good estimate of how ``different'' two routes are.
A different approach could be to discard the current algorithm and instead use
an external dataset of standard routes. In this setting, standard routes need
not be computed. Flights that diverge from them can simply be discovered by
computing the distance from the standard route for the particular pair of
airports, and checking the result against a threshold.
